% Chapter 1

\chapter{Motivations and Problem Statement} % Write in your own chapter title
\label{Chapter2}
\lhead{Chapter 2. \emph{What to Write}} % Write in your own chapter title to set the page header

\section{Motivations}

(add the relevant info related to our project reason open source camera chip - nust , 10x (isp), us (encoder))

% state some industry level uses
<<<<<<< Updated upstream
With the rapid increase in the development of products and benefits offering full-motion digital video, digital video coding is currently gaining importance and has a considerable monetary impact on computer, imaging and telecommunications technology. Throughout the early phases of technology, there have been many international video coding standards as engines behind the commercial success of digital video compression. In comparison to current existing standards, H264 has many new features that makes it more emphatic. The important features that distinguish H264 from other standards are good video quality at high and low bit rates and network friendliness. \cite{amer2005high} 
=======
% ref taken from 0201137
With the rapid increase in the development of products and services offering full-motion digital video, digital video coding is currently gaining importance and has a considerable monetary impact on computer, imaging and telecommunications technology.There have been a lot of efforts to achieve an efficient video encoding format, which offers a better remote viewing quality at a lower bit-rate and uses less bandwidth than the previous standards (i.e., MPEG-2, H.263, or MPEG-4 Part 2). H.264 or MPEG- Part 10 has all these new, better compression features that makes it more emphatic.The primary features that set H.264 apart from competing standards include network friendliness and good video quality at both high and low bit rates.

>>>>>>> Stashed changes

%ref taken from me_jvlsis08
It is proposed that H264 can deliver two times better performance than the previous MPEG-2 coding standard, both in terms of compression efficiency and picture eminence. Moreover, previous H.63 and MPEG-4 implement block-based motion estimation to reduce temporal redundancy between frames. But in H.264, block matching efficiency is improved by some innovative features such as variable block size motion estimation (VBS$\_$ME) and motion vector prediction \cite{li2003serial}. The computational complexity of this model is increased by the factor of four due to these features.


% why are we working on this project
This is a long-term project, in collaboration with 10x Engineers and team from NUST. A team from 10x worked on the ISP part of it whereas our group worked on the encoder part of the camera. This work is unique in Pakistan as very few engineers here work on SoC designs. This makes our project valuable as it can lead to new opportunities for local research and it can make a significant impact on the video compression industry and contribute to the development of new technologies in Pakistan and beyond.
In terms of encoding efficiency and computational complexity, there is still a need for further improvements in the Baseline profile of H.264 especially if we are planning to use this standard for real time applications. Our contribution is a significant step towards enhancing the over-all performance of the CODEC i.e. by optimizing the inter-prediction algorithm to reduce the bit-rate and improve the visual quality of the video. 


